\documentclass{article}
\usepackage{fullpage}
\usepackage{graphicx}

\usepackage[
  letterpaper=true,
  colorlinks=true,
  linkcolor=black]{hyperref}

\newenvironment{icompact}{
  \begin{list}{$\bullet$}{
    \parsep 0pt plus 1pt
    \partopsep 0pt plus 1pt
    \topsep 2pt plus 2pt minus 1pt
    \itemsep 0pt plus 1pt
    \parskip 0pt plus 2pt
    \leftmargin 0.3in}
       \raggedright}
  {\normalsize\end{list}}

\newcounter{ecompactctr}
\newenvironment{ecompact}{
  \begin{list}{\arabic{ecompactctr}. }{
    \usecounter{ecompactctr}
    \parsep 0pt plus 1pt
    \partopsep 0pt plus 1pt
    \topsep 2pt plus 2pt minus 1pt
    \itemsep 0pt plus 1pt
    \parskip 0pt plus 2pt
    \leftmargin 0.3in}
       \raggedright}
  {\normalsize\end{list}}

\renewcommand{\bibitem}[1]{\item }
\newcommand{\comment}[1]{}

\begin{document}

\begin{center}
  {\Large \bf David Samuel Bindel} \\
  Assistant Professor of Computer Science
\end{center}

\begin{trivlist}
\item
Department of Computer Science
  \hfill \href{mail:bindel@cs.cornell.edu}{\tt bindel@cs.cornell.edu} \\
Cornell University
  \hfill \url{www.cs.cornell.edu/~bindel} \\
Ithaca, NY 14853
  \hfill Office: 607-255-5395
%%% Department of Mathematics                  \hfill Office: (212) 998-3155 \\
%%% Courant Institute of Mathematical Sciences \hfill Fax: (212) 995-4121\\
%%% New York University                        \hfill \href{mail:dbindel@cims.nyu.edu}{\texttt{dbindel@cims.nyu.edu}} \\
%%% 251 Mercer Street                          \hfill \href{www.cims.nyu.edu/~dbindel}{\texttt{http://www.cims.nyu.edu/\string~dbindel}} \\
%%% New York, NY 10012
\end{trivlist}


\section*{Education}

\begin{tabular}{ll}
May 1999 &
  B.S. in Mathematics and in Computer Science,
  University of Maryland, College Park \\
December 2006 &
  Ph.D. in Computer Science, University of California, Berkeley
\\
\\
Advisors:
 & James Demmel (Computer Science Division and Department of Mathematics) \\
 & Sanjay Govindjee (Department of Civil Engineering) \\
Dissertation title: &
   {\em Structured and Parameter-Dependent Eigensolvers for
        Simulation-Based Design of} \\
&  {\em Resonant MEMS}
\end{tabular}


\section*{Professional Experience}

\begin{tabular}{ll}
  Summer 2009-present.
    & Assistant Professor, Department of Computer Science,
      Cornell University \\
  Fall 2006-Summer 2009.
    & Courant Instructor of Mathematics, New York University \\
  Fall 1999-Summer 2006.
    & Graduate Student Researcher, CS Division, UC Berkeley \\
  Fall 2005 and Spring 2001.
    & Graduate Student Instructor, CS Division, UC Berkeley
\end{tabular}


\section*{Research interests}

% \begin{icompact}
%   \item Numerical linear algebra, particularly eigenvalue problems
%   \item Modeling microelectromechanical systems (MEMS)
%   \item Computational mechanics and finite element analysis
%   \item Numerical analysis of problems from computer networking and
%     systems
%   \item Social network analysis
%   \item Numerical software engineering
% \end{icompact}

\section*{Awards}

\input{awards}

% ===============================================================

\section*{Publications}

\subsection*{Journal papers}

\input{journal}

\subsection*{Conference papers}

\input{conf}

\subsection*{Reports}

\input{report}

% ---------------------------------------------------------------

\section*{External talks}

\subsection*{Colloquium and seminar talks}

\input{seminar}

\subsection*{Workshop and conference talks}

\input{meeting}

% ---------------------------------------------------------------

\section*{Teaching}

\input{teaching}

% ---------------------------------------------------------------

\section*{PhD students}

These students are all in progress:
\begin{ecompact}
\item
  Eric Lee, Computer Science -- Demand response optimization and fault
  analysis for power grids
\item
  Kun Dong, Applied Math -- Non-negative matrix factorizations for
  topic modeling
\item
  Erdal Yilmaz, Applied and Engineering Physics -- MEMS gyroscope
  simulations (graduating May 2016)
\item
  Amanda Hood, Applied Math -- Nonlinear eigenvalue problems for
  resonances (graduating May 2016)
\item
  Colin Ponce, Computer Science -- Fault diagnosis in electrical networks
  %Community detection in social networks
  (graduating May 2016)
\end{ecompact}

\section*{Masters students}

\begin{ecompact}
\item Batu Inal (2016) -- {\em Super Seminar Scraper}
\item Sania Nagpal (2016) --
  {\em Fast trial spaces for parameterized PageRank problems}
\item Suarabh Netravalkar (2016) --
  {\em Fast trial spaces for parameterized PageRank problems}
\item Shitong Jia (2015) -- {\em Etch-a-Sketch 3D}
\item Jing Jing and Jiankun Lu (2015) -- {\em Super Seminar Scraper}
\item Xiangyu Zhang (2015) -- {\em Edge-weighted parameterized PageRank: a demo}
\item Ziyang Tang (2015) -- {\em Optimizing price of anarchy in a
  model of opinion formation}
\item Jia Zhang (2015) -- {\em Spectral histogram on networks}
\item Xiangyu Zhang and Markus Salasoo (2015) -- {\em Edge-weighted parameterized PageRank: a demo}
\item
  Ting-Hao Chen and Syed Hassan (2013) -- {\em MEMS Simulation as a Service}
\item
  Yong Jae Kim (2013) -- {\em Teaching Support Vector Machines to
    Predict Profitable Stock Offers}
\item
  Jungmin Yun and Jesseon Chang (2012) -- {\em Python implementation
    of a stochastic RBF optimization method}
\item
  Chen-Yi (Charlie) Chen (2012) -- {\em A flexible interface to FEAP}
\item
  Peter Hunt and Scott Purdy (2010) -- {\em Process replication for
    HPC applications}
\item
  Silvio Tarca (2010) -- {\em Performance optimization of symmetric
    factorization algorithms}
\end{ecompact}

\section*{Undergraduate research students}

\begin{ecompact}
  \item Humam Alwassel (2015)
  \item Patrick Chen (2015--2016)
  \item Joo Young Park (2015)
  \item Sheroze Sheriffdeen (2015--2016)
  \item Nimit Sohoni (2015)
  \item Leon Davis (2015)
  \item Greg Rosenthal (2015)
  \item Eric Ma (2015)
  \item Brandon Hartz (2014)
  \item Marshall Jiang (2014--2015)
  \item Kyu-Young Kim (2011--2012)
  \item Jiexun Xu (2008)
  \item Iva Vukicevic (2008)
  \item Daniel Parry (2008)
  \item Anwis Das (2002)
  \item Ernest Zhu (2001)
  \item Wayne Kao (2001)
\end{ecompact}

% ---------------------------------------------------------------

\section*{Released software packages}

\begin{ecompact}
\item {\bf RSC}
  A rank-structured Cholesky factorization code derived from Cholmod.
\item {\bf PySOT}
  A Python package for parallel response-surface based
  surrogate optimization methods.
\item {\bf POAP}
  Python Optimization Asynchronous Pluming.
\item {\bf AxFEM}
  A finite element simulator for computing properties of axisymmetric
  resonant structures such as gyroscopes.
\item {\bf MWwrap}
  A wrapper generator for automating calls to C/C++ and Fortran
  codes from MATLAB.
\item {\bf Matexpr}
  A source-to-source translator for compiling MATLAB-like
  expressions within C/C++.
\item {\bf BoneFEA}
  A commercial code for fast simulation of failure in human bone.
\item {\bf HiQLab}
  A finite element simulator for computing losses in resonant
  microsystems.
\item {\bf MATFEAP/FEAPMEX}
  MATLAB interfaces to the FEAP finite element analysis code.
\item {\bf SUGAR}
  A lumped-element modeling system for MEMS based on modified nodal
  analysis.  I also worked on M\&MEMS, a web-based interface to the
  SUGAR simulator.
\item {\bf CLAPACK}
  C translation of version 3.0 of the LAPACK library for dense numerical
  linear algebra.
\end{ecompact}

% ---------------------------------------------------------------

\section*{Referee activity}

\begin{icompact}
  \item {\em ACM Journal on Emerging Technologies in Computing Systems} (2014)
  \item {\em ACM Transactions on Mathematical Software} (2013, 2014,
    2015, 2016)
  \item {\em American Mathematical Monthly} (2015)
  \item {\em SIAM Journal on Matrix Analysis} (2005, 2006, 2007, 2008,
    2009, 2012, 2014, 2015)
  \item {\em SIAM Journal on Multiscale Modeling and Simulation} (2012)
  \item {\em SIAM Journal on Scientific Computing} (2008, 2009, 2014, 2015)
  \item {\em IEEE Sensors} (2014)
  \item {\em IEEE Transactions on Dependable and Secure Computing} (2014)
  \item {\em IEEE Journal of Microelectromechanical Systems} (2003,
    2007, 2008, 2009, 2010, 2011)
  \item {\em IEEE Transactions on Signal Processing} (2007)
  \item {\em Applicable Analysis} (2007)
  \item {\em Applied Mathematics and Computation} (2011)
  \item {\em Communications in Pure and Applied Math} (2010)
  \item {\em International Journal of Computer Mathematics} (2012)
  \item {\em International Journal of Solids and Structures} (2015)
  \item {\em Journal of Applied Mechanics} (2007)
  \item {\em Journal of Applied Physics} (2007)
  \item {\em Journal of Computational Physics} (2011, 2012, 2015, 2016)
  \item {\em Journal of Micromechanics and Microengineering} (2007,
    2008, 2009, 2011, 2014, 2015)
  \item {\em Journal of Machine Learning Research} (2016)
  \item {\em Journal of Physics A: Mathematical and Theoretical} (2007)
  \item {\em Linear Algebra and its Applications} (2003, 2009, 2010,
    2013, 2014, 2015)
  \item {\em Mathematics of Computation} (2007)
  \item {\em Multiscale Modeling and Simulation} (2013)
  \item {\em Numerical Linear Algebra with Applications} (2010, 2011,
    2012, 2013, 2014, 2015, 2016)
  \item {\em Numerische Mathematik} (2009, 2011)
  \item {\em Transactions of the Canadian Society for Mechanical
    Engineering} (2009)
  \item Technical program commitee, IPDPS 2016.
  \item Technical program commitee, IPDPS 2015.
  \item Technical program commitee, IPDPS 2014.
  \item Posters committee, Supercomputing 2014.
  \item Technical program committee, Supercomputing 2013.
  \item DSL 2011: IFIP Working Conference on Domain-Specific Languages (2011)
  \item PARA04 conference proceedings (2004)
  \item IEEE ARITH conference (2004)
  \item ACM International Conference on Supercomputing (2003)
  \item ACM SIGGRAPH (2009, 2011)
  \item Proposal referee for the Netherlands science foundation (NWO)
  \item Proposal referee for DOE (2009, 2015)
\end{icompact}

% ---------------------------------------------------------------

\section*{Professional activities}

\begin{icompact}
  \item Scientific Co-Chair, SIAM Applied Linear Algebra 2018.
  \item Householder Symposium Committee, 2014--present.
  \item Secretary, SIAM Activity Group on Linear Algebra, January
    2013--January 2016.
  \item Organizer, New York Conference on Applied Mathematics, 2011--2013.
  \item Editor, {\em Numerical Linear Algebra with Applications},
        2010 -- present.
  \item Managing editor,
        {\em Electronic Transactions on Numerical Analysis},
        2008 -- present.
  \item Organizer, Cornell scientific computing and numerics (SCAN) seminar,
        2009 -- present.
  \item Organizer, NYU numerical analysis and scientific computing seminar,
        Fall 2007--Spring 2009.
  \item Secretary, IEEE 754 standard revision committee,
        January 2002 -- March 2004
        (\verb+http://grouper.ieee.org/groups/754/+)
  \item Member: American Mathematical Society,
        Society for Industrial and Applied Mathematics,
        and Association for Computing Machinery
        %, and US Association for Computational Mechanics
\end{icompact}

% ---------------------------------------------------------------

%%\section*{Academic references}

%%\begin{icompact}
%%\item
%%  James Demmel, \\
%%  Computer Science Division,
%%  University of California, Berkeley \\
%%  Berkeley, CA 94720. \\
%%  Phone: (510) 643-5386.
%%  E-mail: \texttt{demmel@eecs.berkeley.edu}.

%%\item
%%  Leslie Greengard, \\
%%  Department of Mathematics,
%%  Courant Institute of Mathematical Sciences,
%%  New York University, \\
%%  251 Mercer Street, \\
%%  New York, NY 10012. \\
%%  Phone: (212) 998-3306.
%%  E-mail: \texttt{greengard@cims.nyu.edu}.

%%\item
%%  Panos Papadopoulos, \\
%%  Department of Mechanical Engineering,
%%  University of California, Berkeley \\
%%  Berkeley, CA 94720. \\
%%  Phone: (510) 642-3358.
%%  E-mail: \texttt{panos@me.berkeley.edu}.

%% \item
%%   Sanjay Govindjee,
%%   Structural Engineering, Mechanics and Materials,
%%   Department of Civil Engineering,
%%   University of California,
%%   Berkeley, CA 94720. \\
%%   Phone: (510) 642-6060.
%%   E-mail: \texttt{sanjay@ce.berkeley.edu}.

%% \item
%%   Roger Howe,
%%   Department of Electrical Engineering,
%%   Integrated Circuits Laboratory,
%%   114 Allen CIS Building,
%%   Stanford University,
%%   Stanford, CA 94305. \\
%%   Phone: (650) 723-8030.
%%   E-mail: \texttt{howe@ee.stanford.edu}.

%% \item
%%   Beresford Parlett,
%%   Department of Mathematics,
%%   University of California,
%%   Berkeley, CA 94720. \\
%%   Phone: (510) 642-6655.
%%   E-mail: \texttt{parlett@math.berkeley.edu}.

%%\end{icompact}


\end{document}
